Software Product Line Engineering via Software Transplantation

### 2024-04-09 paper/FOUNDRY-TOSEM_JOURNAL/sections/sec01-introduction.tex
review rebuttal version
New version of the paper and rebuttal documents generated after Leandro and Justyna reviewes

Software Product Line (SPL) is a systematic methodology for producing related software products from shared development assets~\cite{Pohl2005, Linden2007, Fischer2015}.
%SPL organises a codebase into a core asset features that are shared across all products within the product line. Then, it defines a \emph{dependency relation} over those features and a \emph{variability mechanism} that creates products from subsets of the features, subject to the dependency relation.
SPL organises features as core assets that are shared across all products within the product line. Then, it defines a \emph{dependency relation} over those features and a \emph{variability mechanism} that creates products from subsets of the features, subject to the dependency relation.
%SPL organises a codebase into a \emph{product base}, a set of  features, then defines a \emph{dependency relation} over those features and a \emph{variability mechanism} that creates products from subsets of the features, subject to the dependency relation.
By centralising features into a product base, SPL prevents cross-product feature drift, allowing all products to benefit from a feature's improvement, if they use it.
For companies producing related products, adopting SPL improves productivity and quality, speeds time to market, and reduces cost, because it facilitates the reuse of development artifacts, such as code and design~\cite{Linden2007,Bastos2015}.

Despite its benefits, adopting SPL requires considerable upfront investment before its benefits can be realised. 
The cost of migrating existing products to SPL is lower than adopting SPL from scratch, making \emph{extractive}~\cite{Krueger2001} adoption more common, especially in companies with many software system variants in production~\cite{Berger2013}. Two factors drive this preference: 1) it is often hard to know upfront that SPL will be needed because related products often emerge from a small set of initial products, and 2)  starting from scratch discards considerable knowledge and investment in existing codebases, when they exist~\cite{Breivold2008,Northrop2012}.

@@

%\textcolor{magenta}{From Justyna to Leandro: I am missing here a paragraph that would summarise related work on automating reengineering tasks.}\ls{fixed}

@@

Inspired by this line of work, we introduce \FOUNDRY (\Cref{sec:approach}), the first software transplantation approach for SPL reengineering. \FOUNDRY is independent of the programming language, and supports SPL's \emph{domain engineering} and \emph{application engineering}~\cite{Clements2001} processes at the code level.
It tackles each SPL reengineering task, easing some and automating others. \FOUNDRY does not eliminate the manual labour of feature identification, but reduces it to the task of annotating the entry points (\ie the interface) of a feature, or its ``organ'' using transplantation nomenclature. \FOUNDRY amortises this manual step across a sequence of transplantations. 
\FOUNDRY  automates feature extraction; to do so, it uses slicing to overapproximate feature dependencies.  It leverages transplantation to automate the variability mechanism and, simultaneously, tackle slice-imprecision.
Key to \FOUNDRY is mapping software transplantation's ``over-organs'', conservative program slices~\cite{Barr2015}, to product line assets, or features. 
Its use of slicing means that \FOUNDRY does not need specially prepared donors; The donor programs can even be unaware that they are participating in an SPL. A product line via ST is composed of multiple over-organs and a ``product base'', a host that contains all features are shared across all products within the product line, so constructing a product entails transplanting a set of organs into a product base.

@@

We realise \FOUNDRY in \prodscalpel, a tool that transplants multiple organs (\ie, a set of interesting features) from donor systems into an emergent product line for codebases written in C.  
%\ls{variability mechanism is used without definition, just below:  it needs to be defined with a citation.} 
\prodscalpel also supports the use of existing variability mechanisms~\cite{Gacek2001} based on \emph{feature toggle}~\cite{Rahman2016} or \emph{preprocessor directives}~\cite{Kastner2008B}. It can surround implanted organs with feature flags, which permit enabling and disabling features, to facilitate its integration into an existing SPL codebase that uses them.

>>>>

Software Product Line (SPL) is a systematic methodology for producing related software products from shared development assets~\cite{Pohl2005, Linden2007, Fischer2015}. SPL organises features as core assets that are shared across all products within the product line. Then, it defines a \emph{dependency relation} over those features and a \emph{variability mechanism} that creates products from subsets of the features, subject to the dependency relation. By centralising features into a product base, SPL prevents cross-product feature drift, allowing all products to benefit from a feature's improvement, if they use it.
For companies producing related products, adopting SPL improves productivity and quality, speeds time to market, and reduces cost, because it facilitates the reuse of development artifacts, such as code and design~\cite{Linden2007,Bastos2015}.

Despite its benefits, adopting SPL requires considerable upfront investment before its benefits can be realised \cite{Breivold2008, Bastos2017, Daniel2020}. The cost of migrating existing products to SPL is lower than adopting SPL from scratch, making \emph{extractive}~\cite{Krueger2001} adoption more common, especially in companies with many software system variants in production~\cite{Berger2013}. Two factors drive this preference: 1) it is often hard to know upfront that SPL will be needed because related products often emerge from a small set of initial products, and 2)  starting from scratch discards considerable knowledge and investment in existing codebases, when they exist~\cite{Breivold2008,Northrop2012}.

@@

@@

Inspired by this line of work, we introduce \FOUNDRY (\Cref{sec:approach}), the first software transplantation approach for SPL reengineering. \FOUNDRY is independent of the programming language, and supports SPL's \emph{domain engineering} and \emph{application engineering}~\cite{Clements2001} processes at the code level.

It tackles each SPL reengineering task, easing some and automating others. \FOUNDRY does not eliminate the manual labour of feature identification, but reduces it to the task of annotating the entry points (\ie the interface) of a feature, or its ``organ'' using transplantation nomenclature. \FOUNDRY amortises this manual step across a sequence of transplantations. 
\FOUNDRY  automates feature extraction; to do so, it uses slicing to overapproximate feature dependencies.  It leverages transplantation to automate the variability mechanism and, simultaneously, tackle slice-imprecision.
Key to \FOUNDRY is mapping software transplantation's ``over-organs'', conservative program slices~\cite{Barr2015}, to product line assets, or features. 
Its use of slicing means that \FOUNDRY does not need specially prepared donors; The donor programs can even be unaware that they are participating in an SPL. A product line via ST is composed of multiple over-organs and a ``product base'', a host that contains all features are shared across all products within the product line, so constructing a product entails transplanting a set of organs into a product base.

@@

We realise \FOUNDRY in \prodscalpel, a tool that transplants multiple organs (\ie, a set of interesting features) from donor systems into an emergent product line for codebases written in C. It can be used to make possible the extraction of features from donor systems, adapting them to integrate with the host product base, and ultimately, automating the generation of product lines. \prodscalpel also supports the use of existing variability mechanisms~\cite{Gacek2001} based on \emph{feature toggle}~\cite{Rahman2016} or \emph{preprocessor directives}~\cite{Kastner2008B}. It can surround implanted organs with feature flags, which permit enabling and disabling features, to facilitate its integration into an existing SPL codebase that uses them.

\textcolor{blue}{\FOUNDRY is the conceptual solution for the reengineering of individual systems into SPL, while \prodScalpel is one possible implementation designed to automate the steps outlined by Foundry. As the practical counterpart to \FOUNDRY's conceptual underpinnings, \prodscalpel bridges the gap between theory and implementation.}

### 2024-04-09 paper/FOUNDRY-TOSEM_JOURNAL/sections/sec02-motivating_example.tex
review rebuttal version
New version of the paper and rebuttal documents generated after Leandro and Justyna reviewes

 \begin{figure*}[t]
	\centering \includegraphics[width=\textwidth]{images/incremental_ST4.png}
	\caption{Product derivation process using the \FOUNDRY approach. \textit{\prodscalpel transplants three features, in sequence, into the GEdit's product base to derive a new text editor after three iterations of organ transplantation.} }
	\label{fig:incremental_pd}
\end{figure*} 
 %COMMENT: \ls{The case for SPL needs to be elaborated:  we must argue that the various GNOME subprojects share sufficient code to define a product base.  If too many of them are too different, SPL would be inappropriate.}
%\todo{This sentence is problematic:  we shoudl discuss. Done}
%\eb{from Leandro: I've rewritten this sentence by providing details of how GNOME could benefit from using SPLE principles as mass customisation and a common platform.}

The open-source GNOME project\footnote{https://wiki.gnome.org/Projects} encompasses a large portfolio of individual programs evolve as independently as possible from the rest. These programs share features, but because they are separately developed, their constituent features cannot be easily reused across its portfolio to provide mass customization, at least without much manual effort. The combination of mass customisation and a common platform, principles of Software Product Line Engineering (SPLE)~\cite{Pohl2005}, would allows to GNOME team reuses a common base of technology and, at the same time, to bring out products tailored individual customers. Without a common platform and a software development process base on mass customisation, it may be more difficult to the GNOME project provides customized products and effectively manage the commonality and variability of its features.

The GNOME  project is a natural candidate for SPL, but the significant reengineering investment of time and resources have prevented it from adopting SPL. \FOUNDRY is transformative because it can be used to reduce this cost.

We show how the GNOME team could use \prodscalpel to quickly generate a product line. Suppose project collaborators want to build a product line in the domain of text editors. This product line would allow GNOME to produce text editors that augment its current text editor, \emph{GEdit}\footnote{https://wiki.gnome.org/Apps/Gedit}, with additional features. Since they have decided to augment Gedit, GNOME team would select it as the product base, the shared substrate of a product line that, for \FOUNDRY, serves the host for transplanted features.  Assume that the GNOME team targets the following three features (1) $\texttt{side-panel}$, (2) $\texttt{split pane}$, and (3) $\texttt{presentation}$. They then identify two donors from which to transplant these features:  \emph{NEdit}\footnote{https://sourceforge.net/projects/nedit/}, a multi-purpose text editor that is not part of the GNOME portfolio, and \emph{Evince}\footnote{https://wiki.gnome.org/Apps/Evince}, a document viewer for multiple document formats that is part of GNOME, not not an editor.

\prodscalpel iteratively and incrementally re-engineers the codebase for SPL.  In the donor, engineers need to demarcate all feature entry points to transplant; a single annotation is sufficient for \prodscalpel to extract a feature. 
%\todo{remove: To prepare the host, GNOME engineers use \prodscalpel to extract a product base by removing all not desired features. }

To prepare the host, GNOME engineers use \prodscalpel to extract a product base from an existing system by removing all features not shared across all products within the built product line

Then, the engineers must annotate the host to indicate the implantation point for each target feature, or “organ” using transplantation nomenclature. GNOME engineers then run \prodscalpel on these inputs, once per feature, with 
%\ls{This listing is better.  First, the important thing is to show what a user would \emph{actually} type!  Doing so subtly shows that the system is real.  This text is nothing real and, as such, it invalidates the sentence that proceeds it.  Second, why define these switches?  They're self-defining!  What could the switch "donor\_folder" mean other that the path to the donor source code? I am leaving this todo here briefly for you to read it before I revert your change.  I'm trying to avoid churn.  I would like your help with which ones you think are mandatory.  Many should be either defaulted or in a configuration file, not on the command line.  Thinking more about that, maybe all are configuration, not CLI, and we should present a config file here with a minimal CLI invocation.} \eb{I'm with you Earl, all parameters of this command can be considered configuration. I can improve it in the next version of the ProducScalpel or during the journal review process. Do you prefer I remove this command for the moment?}
\begin{quote}
    $\texttt{./prodScalpel \textemdash \textemdash seeds\_file}$: \textit{ The file which contains the seeds for Genetic Programming(GP) algorithm.}

@@

\Cref{fig:incremental_pd} illustrates all transplantation iterations performed to generate a new product.  It shows a new text editor derived from the transplant of features from different donors and using GEdit as a product base. Using feature models~\cite{Kang1990} to represent each donor system and the product base evolution, \prodscalpel first transplants the \emph{side-panel} feature, extracted from GEdit itself. This transplantation demonstrates that \prodscalpel can transplant features into a product base that comes from the same codebase.  Next, \prodscalpel transplants the \emph{split\_pane} feature from NEdit. It shows how \prodscalpel manages to transplant features from distinct codebases, which is not possible without manual effort using the current state-of-art to SPL reengineering. Finally, \prodscalpel transplants the \emph{presentation} feature from GNOME's Evince renderer.

>>>>

 %COMMENT: \ls{The case for SPL needs to be elaborated:  we must argue that the various GNOME subprojects share sufficient code to define a product base.  If too many of them are too different, SPL would be inappropriate.}
%\todo{This sentence is problematic:  we shoudl discuss. Done}
%\eb{from Leandro: I've rewritten this sentence by providing details of how GNOME could benefit from using SPLE principles as mass customisation and a common platform.}

\textbf{ORIGINAL VERSION}

%to delete
\sout{The open-source GNOME project\footnote{https://wiki.gnome.org/Projects} encompasses a large portfolio of individual programs evolve as independently as possible from the rest. These programs share features, but because they are separately developed, their constituent features cannot be easily reused across its portfolio to provide mass customization, at least without much manual effort. The combination of mass customisation and a common platform, principles of Software Product Line Engineering (SPLE)~\cite{Pohl2005}, would allows to GNOME team reuses a common base of technology and, at the same time, to bring out products tailored individual customers. Without a common platform and a software development process base on mass customisation, it may be more difficult to the GNOME project provides customized products and effectively manage the commonality and variability of its features.}

%\textcolor{blue}{In the dynamic landscape of open-source development, projects such as GNOME \footnote{https://wiki.gnome.org/Projects} harbor an extensive collection of independently evolving programs. While these programs share common features, their separate development trajectories hinder seamless reuse across the entire portfolio for mass customization, necessitating manual efforts. Principles of Software Product Line Engineering (SPLE)\cite{Pohl2005}, embodying mass customization and a common platform, could empower the GNOME team to reuse a foundational technology base while tailoring products to individual customers. However, the adoption of SPLE by GNOME faces a significant obstacle. The intricate web of programs within GNOME, developed independently over time, lacks a common platform and a software development process grounded in mass customization. Consequently, providing customized products and effectively managing the commonality and variability of features becomes a great challenge.}

\textcolor{blue}{In the dynamic landscape of open-source development, projects such as GNOME \footnote{https://wiki.gnome.org/Projects} harbor an extensive collection of independently evolving programs. While these programs share common features, their separate development trajectories hinder seamless reuse across the entire portfolio for mass customization, necessitating manual efforts. Principles of SPLE\cite{Pohl2005}, embodying mass customization and a common platform, could empower the GNOME team to reuse a foundational technology base while tailoring products to individual customers. However, the adoption of SPLE by GNOME faces a significant obstacle. The intricate web of programs within GNOME, developed independently over time, lacks a common platform and a software development process grounded in mass customization. Consequently, providing customized products and effectively managing the commonality and variability of features becomes a great challenge.}
%to delete
\sout{The GNOME  project is a natural candidate for SPL, but the significant reengineering investment of time and resources have prevented it from adopting SPL. \FOUNDRY is transformative because it can be used to reduce this cost.}

\textcolor{blue}{However, the adoption of SPLE by GNOME faces a significant obstacle. The intricate portifolio of programs within GNOME, developed independently over time, lacks a common platform and a software development process grounded in mass customization. Consequently, providing customized products and effectively managing the commonality and variability of features becomes a great challenge.}

%to delete
\sout{We show how the GNOME team could use \prodscalpel to quickly generate a product line. Suppose project collaborators want to build a product line in the domain of text editors. This product line would allow GNOME to produce text editors that augment its current text editor, \emph{GEdit}\footnote{https://wiki.gnome.org/Apps/Gedit}, with additional features. Since they have decided to augment Gedit, GNOME team would select it as the product base, the shared substrate of a product line that, for \FOUNDRY, serves the host for transplanted features. Assume that the GNOME team targets the following three features (1) $\texttt{side-panel}$, (2) $\texttt{split pane}$, and (3) $\texttt{presentation}$. They then identify two donors from which to transplant these features:\emph{NEdit}\footnote{https://sourceforge.net/projects/nedit/}, a multi-purpose text editor that is not part of the GNOME portfolio, and \emph{Evince}\footnote{https://wiki.gnome.org/Apps/Evince}, a document viewer for multiple document formats that is part of GNOME, not not an editor. }

\textcolor{blue}{To elucidate, consider a scenario where the GNOME team aims to swiftly generate a product line within the domain of text editors. This endeavor involves became their current text editor, \emph{GEdit}\footnote{https://wiki.gnome.org/Apps/Gedit}, in a product line with the possiblity to transplant additional features. GEdit is selected as the product base, a shared substrate that, in the context of Foundry, serves as the host for transplanted features.}

\sout{Suppose the team targets three features—(1) $\texttt{side-panel}$, (2) $\texttt{split pane}$, and (3) $\texttt{presentation}$. To enrich GEdit product line with these features, two donors are identified: \emph{NEdit}\footnote{https://sourceforge.net/projects/nedit/}, a multi-purpose text editor external to GNOME, and \emph{Evince}\footnote{https://wiki.gnome.org/Apps/Evince}, a document viewer within GNOME but not an editor.}

\textcolor{blue}{Here's where \FOUNDRY and \prodscalpel} can be used to automate the re-engineering process. \FOUNDRY iteratively and incrementally re-engineers the codebase for SPL. In the donor, engineers need to demarcate all feature entry points to transplant; a single annotation is sufficient for \prodscalpel to extract a feature.

To prepare the host, GNOME engineers use \prodscalpel to extract a product base from an existing system by removing all features not shared across all products within the built product line. Then, the engineers must annotate the host to indicate the implantation point for each target feature, or “organ” using transplantation nomenclature. GNOME engineers then run \prodscalpel on these inputs, once per feature, with

\begin{quote}
    $\texttt{./prodScalpel \textemdash \textemdash seeds\_file}$: \textit{ The file which contains the seeds for Genetic Programming(GP) algorithm.}

@@

 \begin{figure*}[t]
	\centering \includegraphics[width=\textwidth]{images/incremental_ST4.png}
	\caption{Product derivation process using the \FOUNDRY approach. \textit{\prodscalpel transplants three features, in sequence, into the GEdit's product base to derive a new text editor after three iterations of organ transplantation.} }
	\label{fig:incremental_pd}
\end{figure*}

Figure 1 illustrates all transplantation iterations performed to generate a new product. It shows a new text editor derived from the transplant of features from different donors and using GEdit as a product base, using feature models to represent each donor system and the product base evolution. \textcolor{blue}{Importantly, the donor systems in this context are not product lines; we are using feature models to map some of their features, making it easier to understand what has been extracted from the donors' codebase.} \prodscalpel first transplants the \emph{side-panel} feature, extracted from GEdit itself. This transplantation showcases \prodscalpel's ability to transplant features into a product base from the same codebase. Next, \prodscalpel transplants the \emph{split\_pane} feature from NEdit, highlighting its ability to integrate features from diverse sources seamlessly into a unified product base. Finally, \prodscalpel transplants the \emph{presentation} feature from GNOME's Evince renderer, further highlighting its ability to integrate features from diverse sources seamlessly into a unified product base.

\textcolor{blue}{It is imperative to emphasize ProdScalpel's agnostic nature towards the order of transplantations. ProdScalpel is designed to operate independently of the transplantation sequence, providing practitioners with the flexibility to tailor the process according to their specific requirements.  However, it is noteworthy that while ProdScalpel accommodates any transplantation sequence, the choice of order may impact the resulting substrate's compactness by influencing interdependencies between transplanted features. }

### 2024-04-09 paper/FOUNDRY-TOSEM_JOURNAL/sections/sec03-approach.tex
review rebuttal version
New version of the paper and rebuttal documents generated after Leandro and Justyna reviewes

Conceptually, in \FOUNDRY, while the product base provides commonalities (i.e., common features) to the target product line, the variability (i.e., variant features) are provided by the transplantation process, as illustrated in \Cref{fig:product_for_transplantation}. This idea opens new ways for SPLE area by automated construction of different products by transplanting multiple organs into a product base.

\begin{figure}[t]
	\centering \includegraphics[width=\textwidth]{images/product_for_transplantation7.png}
	\centering 
	\caption{An overview of how new products are derived from a product line based on ST.}
	\label{fig:product_for_transplantation}
\end{figure}

@@

\begin{figure}[t]
	\centering  \includegraphics[width=\linewidth]{images/FOUNDRY_DOM6.png}
	\centering \caption{Domain engineering process supported by FOUNDRY. \textit{Four over-organs (A, D, G, L) are extracted from three donor systems and kept in the transplantation platform, with product base consisting of 2 features shared across all products (P and Q).} }
	\label{fig:foundry_dom}
\end{figure}

@@

>>>>

Conceptually, in \FOUNDRY, while the product base provides commonalities (i.e., common features) to the target product line, the variability (i.e., variant features) are provided by the transplantation process. This idea opens new ways for SPLE area by automated construction of different products by transplanting multiple organs into a product base.

\textcolor{blue}{\Cref{fig:product_for_transplantation} illustrates the concept under discussion. In the first part of the figure, we identify three donor systems from which features are transplanted into a codebase. The representation of features B, F, H, and J serves to demonstrate the inherent variability in the product line. These features are extracted into the product line and subsequently implanted into specific products according to their requirements.}

\textcolor{blue}{In the second part of the figure, we have a product base generated from a host system. The features N and O in the host system symbolize elements that need to be removed during the transplantation process due to their irrelevance to the target product line or the resulting products. Features P and Q exemplify the commonality within the product line, as they will be kept in the product base and shared among multiple products. As a result, we have a product line comprising features B, F, H, J, P, and Q, enabling the generation of diverse products through the software transplantation process.}

\begin{figure}[t]
	\centering \includegraphics[width=\textwidth]{images/product_for_transplantation7.png}
	\centering 
	\caption{An overview of how new products are derived from a product line based on ST.\textit{Four over-organs (A, D, G, L) are extracted from three donor systems and kept in the transplantation platform, with product base consisting of 2 features shared across all products (P and Q).}}

	\label{fig:product_for_transplantation}
\end{figure}

@@

\begin{figure}[t]
	\centering  \includegraphics[width=\linewidth]{images/FOUNDRY_DOM6.png}
	\centering \caption{Domain engineering process supported by FOUNDRY.  }
	\label{fig:foundry_dom}
\end{figure}

@@

\FOUNDRY is an approach that combines a conceptual framework with a systematic methodology for the reengineering of individual systems into SPLs. It functions as a process aimed at extracting, adapting, and combining features from various sources to construct a product line. \FOUNDRY is implemented and operationalized through \prodscalpel. While Foundry serves as a conceptual foundation and methodology, \prodscalpel acts as the operational instrument, automating the steps outlined by Foundry. Together, they provide a structured and automated solution for the generation of SPL by transition of features from individual systems to a cohesive product line.

### 2024-04-09 paper/FOUNDRY-TOSEM_JOURNAL/sections/sec04-implementation.tex
review rebuttal version
New version of the paper and rebuttal documents generated after Leandro and Justyna reviewes

\subsection{Automating Vein and Over-Organ Extraction}
As optimisations, \prodscalpel extracts the vein from multiple files. \prodscalpel uses backward slicing to identity each function belonging to the vein and saves it in a copy of its souce file in the transplantation platform. Then, it includes all functions calling existing into vein source code in the \emph{array of over-organ statements} used by GP to achieve an organ from the over-organ~\cite{Barr2015}. Thus,  a function/method in the vein shared between two organs is not duplicated when transplanted.

@@

In the end, \prodscalpel synthesises a call to the extracted organ to execute and test it from the wrapper constructed.

>>>>

\textcolor{blue}{In the context of the donor preparation for the transplant, the choice between retaining or removing a portion of code surrounded by directives in the donor depends on the relevance of the code to the target organ. If the portion of code is used as a substrate for the organ to work correctly, ProdScalpel retains the code in the donor, removing only its directives. Otherwise, it removes both the code and the directives that delimit it. Thus, the process illustrated in \Cref{fig:reconfigurator} results in a clean donor codebase, important for providing an organ free of directives for the transplantation process.}

\subsection{Automating Vein and Over-Organ Extraction}
As optimizations, \prodscalpel extracts the vein from multiple files. \prodscalpel uses backward slicing to identity each function belonging to the vein and saves it in a copy of its source file in the transplantation platform. Then, it includes all functions calling existing into vein source code in the \emph{array of over-organ statements} used by GP to achieve an organ from the over-organ~\cite{Barr2015}. Thus,  a function/method in the vein shared between two organs is not duplicated when transplanted.

@@

In the end, \prodscalpel synthesises a call to the extracted organ to execute and test it from the wrapper constructed.
\textcolor{blue}{The organ-host wrapper can be designed to be versatile, enabling the calling of new features as different programs or executable in various languages. This flexibility is a deliberate design choice, aiming to accommodate diverse features from different sources and facilitate their integration into the product line.
Adapting the process to such a polyglot product line can introduce a level of indirection that may have performance implications, especially if external processes like "exec" are involved. 
Additionally, this would have direct impact on the GP algorithm that must be redesigned to handle different languages or execution environments.}

### 2024-04-09 paper/FOUNDRY-TOSEM_JOURNAL/sections/sec06-case_studies.tex
review rebuttal version
New version of the paper and rebuttal documents generated after Leandro and Justyna reviewes

\begin{quote}
\textbf{RQ2.} \emph{Which features can \prodscalpel~productize?}
\end{quote}
We discuss the characteristics of features that \prodscalpel~can and those it cannot yet handle. Given the inherent complexity of the process, it is important to clearly define its current limitations.

@@

\textbf{Subject systems}. We chose two text editors, \emph{VI} and \emph{VIM}, as product bases.
For donors we chose subjects from two different domains: code analysis software and text editors.
We chose \emph{GNU Cflow}, a call graph extractor from C source code, as a donor.
This is to show that donors can also come from different application domains.
We reuse \emph{VI} and \emph{VIM} as donors, as well as another text editor, \emph{kilo}.
This is to show that donors can come from the same system as the product base, but also different systems within the same application domain.

Our donors are presented in Table~\ref{tab:donors_list}.
We identified the following features as possible desired features in a new editor: $\texttt{output}$ from CFLOW, $\texttt{enableRawMode}$ from kilo, $\texttt{vclear}$ from VI, and $\texttt{spell\_check}$ and $\texttt{search}$ from VIM. Table~\ref{tab:donors_list} presents more details on the subjects.

@@

@@

The relatively small number and diversity of systems used for transplantation pose an external threat to validity. However, we tried to mitigate it by constructing a possible real-world scenario, i.e., transplantation of features that would be useful in production of new text editors.

>>>>

\begin{quote}
\textbf{RQ2.} \emph{\textcolor{blue}{How effective is \prodscalpel~at productizing software?}}
\end{quote}
\textcolor{blue}{We conduct an initial empirical study and} discuss the characteristics of features that \prodscalpel~can and those it cannot yet handle. Given the inherent complexity of the process, it is important to clearly define its current limitations.

@@

\sout{\textbf{Subject systems}. We chose two text editors, \emph{VI} and \emph{VIM}, as product bases.
For donors we chose subjects from two different domains: code analysis software and text editors.
We chose \emph{GNU Cflow}, a call graph extractor from C source code, as a donor.
This is to show that donors can also come from different application domains.
We reuse \emph{VI} and \emph{VIM} as donors, as well as another text editor, \emph{kilo}.
This is to show that donors can come from the same system as the product base, but also different systems within the same application domain.}

\textcolor{blue}{\textbf{Subject Systems and Selection Criteria.}To define the context for performing the case study research, the first step was to identify the most suitable codebases for composing the product lines. The selection of subject systems for our case studies was based on four criteria aimed at providing a comprehensive evaluation of the proposed approach and tool. These selection criteria were based on the convenient sampling method Wohlin et al.~\cite{Wohlin2012}.}

\textcolor{blue}{\textbf{(i) Diversity and variation:} The selection of subjects should aim to encompass a range of diversity and variation to provide a comprehensive understanding of the phenomenon under investigation. This can include diversity in terms of domains, sizes, complexity, functionality, or other relevant characteristics.}

\textcolor{blue}{\textbf{(ii) Representativeness:} The selected subjects should be representative of the broader population or the specific context under study. They should capture the characteristics and variations found in the target population or application domain.}

\textcolor{blue}{\textbf{(iii) Feasibility and accessibility:} Practical considerations, such as availability of data, availability of use and modification, access to organizations or systems, and resources required for data collection, should be taken into account during subject selection.}

\textcolor{blue}{\textbf{(iv) Established systems:} Established systems are typically used in real-world scenarios and have a significant user base. Studying such systems can provide insights into practical challenges and solutions.}

\textcolor{blue}{Base on these criteria, we present the systems used in Table~\ref{tab:donors_list}. We chose two text editors, \emph{VI} and \emph{VIM}, as product bases. These text editors serve as the foundation upon which the transplantation process is performed. By selecting widely used and established text editors, we increase the relevance of your case study.}

\textcolor{blue}{To illustrate that donors can come from different application domains, we included \textit{GNU Cflow} one of the donors. \textit{GNU Cflow}, being a call graph extractor from C source code, represents a distinct domain compared to the text editors. This demonstrates that the transplantation process can be applied to diverse types of software systems. In addition to using donors from different domains, we also included \textit{VI} and \textit{VIM} as donors as well as another text editor, \emph{kilo}. This highlights the possibility of using donors from the same system as the product base or from different systems within the same application domain. It showcases the flexibility of the transplantation process and its ability to handle variations within a specific domain.}

We identified the following features as possible desired features in a new editor: $\texttt{output}$ from CFLOW, $\texttt{enableRawMode}$ from kilo, $\texttt{vclear}$ from VI, and $\texttt{spell\_check}$ and $\texttt{search}$ from VIM. Table~\ref{tab:donors_list} presents more details on the subjects.

@@

\textcolor{blue}{In the transplantation of features from two different versions of the same system - one with the desired feature as the donor and another without it as the host - \prodscalpel replicated the original code, incorporating the transplanted feature. The introduction of the organ-host wrapper, a necessary addition, did not compromise the overall integrity or functionality of the codebase. To thoroughly validate this integration, acceptance tests were conducted on the resulting code. These tests affirmed the successful assimilation of the transplanted organ into the original codebase.}

\textcolor{blue}{The capability to selectively remove or reimplant features in a software product line, facilitated by tools like \prodScalpel, can provide adaptability to the product line, allowing dynamic adjustments to meet changing requirements or market demands, providing flexibility in tailoring product configurations without extensive manual efforts. In terms of maintenance and evolution, the process of reimplanting features can automate updates or modifications, replacing intricate manual interventions. The systematic reuse of features across different product variants reduces redundancy and promotes consistency. This removal and reimplantation functionality contributes to product line customization, facilitating tailored solutions for specific customer needs. Additionally, the automated process significantly reduces development time by minimizing manual efforts, ensuring a more efficient product line engineering process. Furthermore, it enhances consistency and cohesion by uniformly applying changes to features across product variants, mitigating the risk of inconsistencies or errors.}

@@

The relatively small number and diversity of systems used for transplantation pose an external threat to validity. However, we tried to mitigate it by constructing a possible real-world scenario, i.e., transplantation of features that would be useful in production of new text editors. By selecting different subject systems allowed us to assess the effectiveness, applicability, and versatility of \prodscalpel in achieving products. We used subjects from different domains and with a wide range of sizes to give evidence that \prodscalpel can also be used to achieve a product line from a different set of usage scenarios.
